Reanalysis Step 4: Main Effects Outcome Analysis

### 2025-08-27 reanalysis-4-outcome_analysis.qmd
debug mi

# ANALYSIS MODE: Use same as imputation
analysis_mode <- "final"  # Should match imputation mode

@@

```{r pool-results}
# Extract coefficient for lapse effect from each imputation
lapse_coefficients <- sapply(model_summaries, function(x) {
  x$estimate[x$term == "ever_lapse_binary"]
})

lapse_ses <- sapply(model_summaries, function(x) {
  x$std.error[x$term == "ever_lapse_binary"]
})

# Pool using Rubin's rules (via mice)
# Create a mira object for pooling
model_list <- lapply(fitted_models, as.mira)
class(model_list) <- c("mira", "list")

# Pool the results
pooled <- pool(model_list)
pooled_results <- summary(pooled, conf.int = TRUE) |> 
  as.data.frame()

@@

```{r marginal-effects}
# Calculate AME for the main effect of lapse
# Use the first few imputations for AME calculation
n_ame_imps <- min(10, n_imputations)  # Use up to 10 imputations for AME

ame_results <- list()
for (i in 1:n_ame_imps) {
  imp_data <- imputed_datasets[[i]]
  imp_data$weights <- twang_weights[[i]]

  # Create survey design to properly handle clustering
  design <- svydesign(
    ids = ~ e_mrn_deidentified,  # Cluster on patient ID
    weights = ~ weights,
    data = imp_data
  )

  # Calculate AME with survey design
  # Note: marginaleffects handles survey objects appropriately
  ame <- avg_slopes(
    fitted_models[[i]], 
    variables = "ever_lapse_binary",
    newdata = imp_data,  # Use the full dataset
    vcov = ~e_mrn_deidentified,  # Cluster standard errors on patient ID
    wts = imp_data$weights  # Apply propensity score weights
  )

  ame_results[[i]] <- data.frame(
    imp = i,
    ame = ame$estimate,
    se = ame$std.error
  )
}

# Combine AME results
ame_combined <- do.call(rbind, ame_results)

# Pool AME estimates using Rubin's rules
pooled_ame <- data.frame(
  Effect = "Average Marginal Effect of Lapse",
  AME = mean(ame_combined$ame),
  SE = sqrt(mean(ame_combined$se^2) + var(ame_combined$ame) * (1 + 1/n_ame_imps)),
  Interpretation = paste0(
    "On average, lapsing from care increases the probability of VI/blindness by ",
    round(mean(ame_combined$ame), 3)
  )
)

@@

>>>>

# ANALYSIS MODE: Use same as imputation
analysis_mode <- "test"  # Should match imputation mode

@@

```{r pool-results}
# Create a proper mira object for mice pooling
# The key is to create a list with the correct structure
mira_object <- list(
  call = fitted_models[[1]]$call,
  call1 = fitted_models[[1]]$call,
  nmis = rep(1, n_imputations),  # Placeholder
  analyses = fitted_models
)
class(mira_object) <- "mira"

# Now pool using mice's native function
pooled <- pool(mira_object)

# Get pooled results with confidence intervals
pooled_results <- summary(pooled, conf.int = TRUE) |> 
  as.data.frame()

@@

```{r marginal-effects}
# Calculate AME for each imputed dataset then pool
ame_list <- list()

for (i in 1:n_imputations) {
  imp_data <- imputed_datasets[[i]]
  imp_data$weights <- twang_weights[[i]]

  # Calculate AME with clustering and weights
  # marginaleffects handles survey designs well
  ame <- avg_slopes(
    fitted_models[[i]], 
    variables = "ever_lapse_binary",
    newdata = imp_data,
    vcov = ~e_mrn_deidentified,  # Cluster SEs on patient ID
    wts = imp_data$weights  # Apply PS weights
  )

  ame_list[[i]] <- ame
}

# Extract estimates and SEs for pooling
ame_estimates <- sapply(ame_list, function(x) x$estimate)
ame_ses <- sapply(ame_list, function(x) x$std.error)

# Apply Rubin's rules for pooling
pooled_ame_est <- mean(ame_estimates)
within_var <- mean(ame_ses^2)
between_var <- var(ame_estimates)
total_var <- within_var + between_var * (1 + 1/n_imputations)
pooled_ame_se <- sqrt(total_var)

# Calculate FMI and df
fmi <- (between_var * (1 + 1/n_imputations)) / total_var
df_ame <- (n_imputations - 1) / fmi^2

# CI using t-distribution
t_crit <- qt(0.975, df_ame)
ci_lower <- pooled_ame_est - t_crit * pooled_ame_se
ci_upper <- pooled_ame_est + t_crit * pooled_ame_se

pooled_ame <- data.frame(
  Effect = "Average Marginal Effect of Lapse",
  AME = pooled_ame_est,
  SE = pooled_ame_se,
  CI_Lower = ci_lower,
  CI_Upper = ci_upper,
  FMI = fmi,
  Interpretation = paste0(
    "On average, lapsing from care increases the probability of VI/blindness by ",
    round(pooled_ame_est, 3)
  )
)

@@

## Summary

The main effects model shows:

- **Primary finding**: Lapsing from care has a significant effect on vision impairment/blindness (OR = `r round(lapse_effect$OR, 2)`, p `r ifelse(lapse_effect$p.value < 0.001, "< 0.001", paste("=", round(lapse_effect$p.value, 3)))`)
- **Effect size**: The average marginal effect indicates that lapsing increases the probability of VI/blindness by `r round(pooled_ame_est, 3)`
